---
title: "Visualizing Data with ggplot"
execute:
  echo: true
subtitle: "ISI-BUDS 2025"
format: 
  revealjs:
    slide-number: true
    incremental: true
    theme: ["../slide-style.scss"]
    title-slide-attributes: 
      data-background-size: 12%
      data-background-position: 50% 75%
---

```{r}
#| echo: false
#| message: false
#| warning: false
library(openintro)
library(tidyverse)
library(titanic)
library(janitor)

theme_set(theme_gray(base_size = 18))

titanic <- titanic_train %>% 
  clean_names() %>% 
  select(survived, pclass, sex, age, fare, embarked) %>% 
  mutate(
    pclass = case_when(
      pclass == 1 ~ "First", 
      pclass == 2 ~ "Second",
      pclass == 3 ~ "Third"
    ),
    embarked = case_when(
      embarked == "S" ~ "Southampton",
      embarked == "C" ~ "Cherbourg",
      embarked == "Q" ~ "Queenstown"
    ),
    embarked = as.factor(embarked),
    sex = as.factor(sex),
    survived = as.logical(survived)
  )
```

## Data Visualizations

- are graphical representations of data


- use different colors, shapes, and the coordinate system to summarize data


- tell a story


- are useful for exploring data


## Data

```{r}
#| echo: true
head(titanic)
```

## Data

```{r}
#| echo: true
tail(titanic)
```

## Data

```{r}
#| echo: true
glimpse(titanic)
```


##

::: {.font75}

Visuals with a Single Categorical Variable

:::


:::: {.columns}

::: {.column width="40%"}

If you could speak to R in English, how would you tell R to make this plot for you?

OR

If you had the data and had to draw this bar plot by hand, what would you do?

:::

::: {.column width="60%"}

```{r echo=FALSE}
ggplot(titanic, aes(x = pclass)) +
  geom_bar()
```

:::

::::





## 3 Steps of Making a Basic ggplot

1. Pick data

2. Map data onto aesthetics


3. Add the geometric layer



## Step 1 - Pick Data

:::: {.columns}

::: {.column width="50%"}

```{r}
#| eval: false
ggplot(
  data = titanic
)
```

:::


::: {.column width="50%"}

```{r}
#| echo: false
#| fig.height: 5
ggplot(titanic)
```

:::

::::

## Step 2 - Map Data to Aesthetics

:::: {.columns}

::: {.column width="50%"}

```{r `code-line-numbers`="3"}
#| eval: false
ggplot(
  data = titanic,
  aes(x = pclass)
) 
```

:::



::: {.column width="50%"}

```{r}
#| echo: false
#| fig.height: 5
ggplot(
  data = titanic,
  aes(x = pclass)
)
```

:::

::::



## Step 3 - Add the Geometric Layer

:::: {.columns}

::: {.column width="50%"}

```{r `code-line-numbers`="5"}
#| eval: false
ggplot(
  data = titanic,
  aes(x = pclass)
) +
  geom_bar() 
```

:::

::: {.column width="50%"}

```{r}
#| echo: false
#| fig.height: 5
ggplot(
  data = titanic,
  aes(x = pclass)
) +
  geom_bar()
```


:::

::::

## Visualizing a Single Categorical Variable with a Bar Plot

::: {.panel-tabset}

## Plot

```{r}
#| echo: false
#| fig-height: 5
#| fig-align: 'center'
ggplot(
  data = titanic,
  aes(x = pclass)
) +
  geom_bar()
```

## English

Create a ggplot using the `titanic` data frame.

Map the `pclass` to the x-axis. 

Add a layer of a bar plot.

## R

```{r}
#| eval: false
#| fig-height: 5
ggplot(
  data = titanic,
  aes(x = pclass)
) +
  geom_bar()
```


:::


## Visualizing a Single Numeric Variable

## Density

::: {.panel-tabset}

## Plot

```{r}
#| echo: false
#| fig-height: 4
#| fig-align: 'center'
ggplot(
  data = titanic,
  aes(x = fare)
) +
  geom_density() 
```


## English

Create a ggplot using the `titanic` data frame.

Map the `fare` to the x-axis. 

Add a layer of a density.


## R

```{r}
#| eval: false
#| fig-height: 2
ggplot(
  data = titanic,
  aes(x = fare)
) +
  geom_density() 
```

:::

## Histogram

::: {.panel-tabset}

## Plot

```{r}
#| echo: false
#| fig-height: 4
#| fig-align: 'center'
ggplot(
  data = titanic,
  aes(x = fare)
) +
  geom_histogram() 
```


## English

Create a ggplot using the `titanic` data frame.

Map the `fare` to the x-axis. 

Add a layer of a histogram.


## R

```{r}
#| eval: false
#| fig-height: 2
ggplot(
  data = titanic,
  aes(x = fare)
) +
  geom_histogram() 
```

:::

## Boxplot

::: {.panel-tabset}

## Plot

```{r}
#| echo: false
#| fig-height: 4
#| fig-align: 'center'
ggplot(
  data = titanic,
  aes(x = fare)
) +
  geom_boxplot() 
```


## English

Create a ggplot using the `titanic` data frame.

Map the `fare` to the x-axis. 

Add a layer of a boxplot


## R

```{r}
#| eval: false
#| fig-height: 2
ggplot(
  data = titanic,
  aes(x = fare)
) +
  geom_boxplot() 
```

:::

<!-- ## Etymology -->


<!-- __histo__ comes from the Greek word _histos_ that literally means "anything set up right". -->

<!-- __gram__: comes from the Greek word _gramma_ which means "that which is drawn". -->

<!-- [Online Etymology Dictionary] -->

<!-- ## Histogram bins -->

<!-- ```{r} -->
<!-- #| echo: false -->
<!-- #| fig-height: 5 -->
<!-- #| fig-align: 'center' -->
<!-- ggplot( -->
<!--   data = titanic, -->
<!--   aes(x = fare) -->
<!-- ) + -->
<!--   geom_histogram() -->
<!-- ``` -->


<!-- ## -->


<!-- ```{r `code-line-numbers`="5"} -->
<!-- #| fig.height: 5 -->
<!-- #| fig.align: 'center' -->
<!-- ggplot( -->
<!--   data = titanic, -->
<!--   aes(x = fare) -->
<!-- ) + -->
<!--   geom_histogram(binwidth = 15)  -->
<!-- ``` -->


<!-- ::: -->

<!-- ::: {.panel-tabset} -->


<!-- ## binwidth = 15 -->



<!-- ```{r echo = FALSE, fig.height=5, fig.align='center'} -->
<!-- ggplot( -->
<!--   data = titanic, -->
<!--   aes(x = fare) -->
<!-- ) + -->
<!--   geom_histogram(binwidth = 15) -->
<!-- ``` -->

<!-- ## binwidth = 50 -->

<!-- ```{r echo = FALSE, fig.height=5, fig.align='center'} -->
<!-- ggplot( -->
<!--   data = titanic, -->
<!--   aes(x = fare) -->
<!-- ) + -->
<!--   geom_histogram(binwidth = 50) -->
<!-- ``` -->

<!-- ## binwidth = 100 -->

<!-- ```{r echo = FALSE, fig.height=5, fig.align='center'} -->
<!-- ggplot( -->
<!--   data = titanic, -->
<!--   aes(x = fare) -->
<!-- ) + -->
<!--   geom_histogram(binwidth = 100) -->
<!-- ``` -->

<!-- ::: -->

<!-- ##  -->

<!-- [There is no "best" number of bins](https://en.wikipedia.org/wiki/Histogram#Number_of_bins_and_width) -->




<!-- ## -->

<!-- :::.{font150} -->

<!-- `r emojifont::emoji(emojifont::search_emoji('rainbow'))[1]` -->
<!-- ::: -->


<!-- Pick your favorite color(s) from the list at: -->

<!-- [bit.ly/colors-r](https://bit.ly/colors-r) -->

<!-- or you can use hex codes. -->

<!-- ## -->

<!-- ```{r `code-line-numbers`="4"} -->
<!-- #| fig-align: 'center' -->
<!-- #| fig-height: 5 -->
<!-- ggplot(data = titanic, -->
<!--        aes(x = fare)) + -->
<!--   geom_histogram(binwidth = 15,  -->
<!--                  color = "white")  -->
<!-- ``` -->

<!-- --- -->

<!-- ```{r `code-line-numbers`="5"} -->
<!-- #| fig-align: 'center' -->
<!-- #| fig-height: 5 -->
<!-- ggplot(data = titanic, -->
<!--        aes(x = fare)) + -->
<!--   geom_histogram(binwidth = 15,  -->
<!--                  fill = "darkred")  -->
<!-- ``` -->

<!-- --- -->

<!-- ```{r} -->
<!-- #| fig-align: 'center' -->
<!-- #| fig-height: 5 -->
<!-- ggplot(data = titanic, -->
<!--        aes(x = fare)) + -->
<!--   geom_histogram(binwidth = 15,  -->
<!--                  color = "white",  -->
<!--                  fill = "darkred")  -->
<!-- ``` -->
